### 2022-03-29 Generic Species.Rmd
New repository with C++ changes

---
  title: "Modeling 3 species on George's Bank (Yellowtail Flounder, Cod, Haddock)"
editor_options: 
  chunk_output_type: console
---

First read in habitat created using point data and covariates (sediment and depth)
and create other required info

#LOAD HABITAT PREVIOUSLY CREATED
hab <- readRDS(file="hab_justYT.RDS")

#plot habitat
par(mar=c(1,1,1,1))
fields::image.plot(hab[["hab"]][["spp1"]])

#plot spawn habitat
par(mar=c(1,1,1,1))
fields::image.plot(hab[["spwn_hab"]][["spp1"]])

@@

#NEW VERSION that has week breaks for entire simulation and allows fishing on just 1 day per week
source("R/init_sim_Bens_nofish.R")
sim <- init_sim_Bens_nofish(nrows = nrow(hab$hab$spp1), ncols = ncol(hab$hab$spp1), n_years = 22, n_tows_day = 1,n_days_wk_fished = 1, n_fleets = 1, n_vessels = 1, n_species = 1,
                            move_freq = 1)

@@

@@

 source("R/init_pop_Bens.R")
# profvis({

#decreasing population settings
Pop <- init_pop_Bens(sim_init = sim, Bio = c("spp1" = 3194), #CHANGE
                     hab = hab[["hab"]], start_cell = c(42,155),
                     lambda = c("spp1" = 0), #set at 0 first, change after initial population set
                     init_move_steps = 20,
                     rec_params = list("spp1" = c("model" = "BH", "a" = 2, "b" = 4, "cv" = 0)), #CHANGE
                     rec_wk = list("spp1" = 9:12),
                     spwn_wk = list("spp1" = 9:12), #CORRECT FOR YTF
                     M = c("spp1" = 0.275), #CHANGE
                     K = c("spp1" = 0.3, "spp2" = 0.3), #CHANGE
                     nz = nonzero_idx
)

#})

#set lambda at 0 first, change after initial population set
Pop$dem_params$spp1$lambda <- .1

@@

fields::image.plot(Pop$Start_pop[[1]], main = "spp1 starting biomass")

@@

```{r temp}
#load temp gradient

moveCov <- readRDS(file="20 year moveCov matrices/GeorgesBank/GB_22yr_ConstTemp_HaddockStrata")

#set temperature preferences manually. 
#The following assumes moveCov has been created and already has an empty spp_tol sublist
moveCov[["spp_tol"]] <- list() #just in case
moveCov[["spp_tol"]] <- list("spp1" = list("mu" = 9, "va" = 4),  #Haddock
     "spp2" = list("mu" = 8.75, "va" = 4.25),  #Cod
     "spp3" = list("mu" = 9, "va" = 4) )    #Yellowtail

@@

```{r fleets}

library(MixFishSim)

@@

@@

@@

  #this starts the profiling
library(profvis)

profvis({

start_time <- Sys.time() # record start time

@@

end_time <- Sys.time() # record end time
tot_time <- end_time - start_time # calculate duration of lrren() example

>>>>

---
  title: "Simple MixFishSim Example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simple MixFishSim Example}
%\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

  This is a simple example of how to use \textbf{'MixFishSim'} to generate
simulations of the dynamics in a mixed fishery. We describe how to calibrate
the habitat fields, the population models, the fishery model and implement a
simple fixed spatial closure. \\

First, load the packages and set a seed for reproducibility.

# Load MixFishSim

```{r packages}
#install.packages('devtools')
#install.packages('githubinstall')

# library(devtools)
# library(githubinstall)
# install_github("pdolder/MixFishSim")
# 
# 
# githubinstall("MixFishSim")

#install.packages('Rcpp')
#install.packages('Rtools')
#install.packages('MixFishSim')
library(MixFishSim)
#library(knitr)
#opts_chunk$set(tidy = TRUE)

#library(reshape2)

#set.seed(123)

@@

#NEW VERSION that has week breaks for entire simulation and allows fishing on just 1 day per week
source("R/init_sim_Bens_nofish.R")
sim <- init_sim_Bens_nofish(nrows = 100, ncols = 100, n_years = 1, n_tows_day = 1,
                            n_days_wk_fished = 1, n_fleets = 1, n_vessels = 1, n_species = 2,
                            move_freq = 1)

@@

This function creates the spatial fields which support the fish populations and
determine their spatial distributions. You define the parameters for the
matern covariance function for each population and optionally the location of
any spawning closure areas.

source("R/BENS_plot_habitat.R")

#values settled on from anisotropy and habtest scripts

spp.ctrl = list(
  "spp.1" = list('nu' = 1/0.05, 
                 'var' = 1,
                 'scale' = 5, 
                 'Aniso' = matrix(nc = 2,  c(1.5, -3, 3, 4) )),
  "spp.2" = list('nu' = 1/0.015, 
                 'var'  = 1,
                 'scale' = 25, 
                 'Aniso' = matrix(nc = 2,c(1, -2, 1, 2))),
  plot.dist = TRUE, 
  plot.file = "testfolder"
)

#DEFINING UNIFORM NXN GRID
#defining strata coordinates
#number of rows and columns to define (assume want n equal sized strata)
nrows <- 4
ncols <- 4

#of the form  c(x1, x2, y1, y2) THESE ARE BOUNDARIES OF STRATA VALUES
# xs are rows ys columns
strata_num <- 1
stratas <- list()

 for(r in seq(nrows)){

   for(c in seq(ncols)){

   stratas[[paste("strata",strata_num,sep="")]] = c( (r-1)*(sim$idx[["nrows"]]/nrows)+1, (r)*(sim$idx[["nrows"]]/nrows) , (c-1)*(sim$idx[["ncols"]]/ncols)+1  , (c)*(sim$idx[["ncols"]]/ncols) )

    strata_num <- strata_num +1

#DEFINING RANDOM NXN GRID
#size of our domain
totalrows <- 100
totalcols <- 100

#desired dimentions of strata
nrows <- 5
ncols <- 6

#of the form  c(x1, x2, y1, y2) THESE ARE BOUNDARIES OF STRATA VALUES
# xs are rows ys columns
strata_num <- 1
stratas <- list()

#generate randow sequence of rows
rowind <- sort(sample(seq(totalrows),nrows-1,replace=FALSE))
rowind <- c(1,rowind,totalrows)

row_idx <- 0
 for(r in seq(length(rowind)-1)){
row_idx <- row_idx+1

   #generate random sequence of columns
    colind <- sort(sample(seq(totalcols),ncols-1,replace=FALSE))
    colind <- c(1,colind,totalcols)

    col_idx <- 0

   for(c in seq((length(colind)-1))){

     col_idx <- col_idx+1

   stratas[[paste("strata",strata_num,sep="")]] = c( rowind[row_idx] , rowind[row_idx+1] , colind[col_idx] , colind[col_idx+1])

    strata_num <- strata_num +1

source("R/BENS_create_hab.R")

hab <- BENS_create_hab(sim_init = sim, 
                       spp.ctrl = spp.ctrl,

                       spawn_areas = list(
                         "spp1" = list(
                           'area1' = c(30,45,55,65),   #of the form  c(x1, x2, y1, y2) THESE ARE BOUNDARIES OF MATRIX VALUES
                           'area2' = c(70,90,50,60)   #need to revisit closure areas
                         ),
                         "spp2" = list(
                           'area1' = c(30,45,55,65),
                           'area2' = c(70,90,50,60)
                         )),
                       spwn_mult = 10,#THIS DOES NOT CHANGE ANYTHING. MUST CHANGE VALUE AT TOP OF BENS_CREATE_HAB 
                       plot.dist = F, 
                       plot.file = "testfolder",

                       #created this new part defining strata
                      strata = stratas

                       )

#print(hab)
source("R/BENS_plot_habitat.R")
## Plot the unadjusted habitat fields
BENS_plot_habitat(hab = hab$hab, 	spp.ctrl = spp.ctrl)

#old version
plot_habitat(hab$hab)

## Plot the adjusted habitat fields
plot_habitat(hab$spwn_hab)

@@

#Week 13 is first week of april
#half way through april to half way through may would be week 15-18

source("R/init_pop_Bens.R")

#original settings
# Pop <- init_pop_Bens(sim_init = sim, Bio = c("spp1" = 2e5, "spp2" = 4e5), #these values from paper : 1e5 and 2e5
# 		hab = hab[["hab"]], start_cell = c(50,50), 
# 		lambda = c("spp1" = 0.1, "spp2" = 0.1), #same lambda for all?
# 		init_move_steps = 20,
# 		rec_params = list("spp1" = c("model" = "BH", "a" = 6, "b" = 4, "cv" = 0.7),
# 				  "spp2" = c("model" = "BH", "a" = 27, "b" = 4,"cv" = 0.6)), #these values from paper 
# 		rec_wk = list("spp1" = 12:15, "spp2" = 12:15),
# 		spwn_wk = list("spp1" = 15:18, "spp2" = 15:18),
# 		M = c("spp1" = 0.2, "spp2" = 0.1), #these values from paper: c("spp1" = 0.2, "spp2" = 0.1)
# 		K = c("spp1" = 0.3, "spp2" = 0.3) #all the same for now
# 		)

#decreasing population settings
Pop <- init_pop_Bens(sim_init = sim, Bio = c("spp1" = 4e5, "spp2" = 10e5), #these values from paper : 1e5 and 2e5
                     hab = hab[["hab"]], start_cell = c(50,50),
                     lambda = c("spp1" = 0.1, "spp2" = 0.1), #same lambda for all?
                     init_move_steps = 20,
                     rec_params = list("spp1" = c("model" = "BH", "a" = 2, "b" = 4, "cv" = 0),
                                       "spp2" = c("model" = "BH", "a" = 7, "b" = 4,"cv" = 0)), #these values from paper
                     rec_wk = list("spp1" = 15:18, "spp2" = 15:18),
                     spwn_wk = list("spp1" = 15:18, "spp2" = 15:18),
                     M = c("spp1" = 0.275, "spp2" = 0.225), #these values from paper: c("spp1" = 0.2, "spp2" = 0.1)
                     K = c("spp1" = 0.3, "spp2" = 0.3), #all the same for now
                     nz = nonzero_idx
                     )

@@

names(Pop)

Pop$dem_params

par(mfrow = c(2,1))

image(Pop$Start_pop[[1]], main = "spp1 starting biomass")
image(Pop$Start_pop[[2]], main = "spp2 starting biomass")

@@

```{r temp}

#set temperature preferences manually. 
#The following assumes moveCov has been created and already has an empty spp_tol sublist
moveCov[["spp_tol"]] <- list() #just in case
moveCov[["spp_tol"]] <- list("spp1" = list("mu" = 7.98, "va" = 3),  #8.13 IF TEMP INCREASES   7.98 if temp constant
                             "spp2" = list("mu" = 7.98, "va" = 3) )

plot(norm_fun(x = 0:25, mu = 8.3, va = 3)/max(norm_fun(0:25, 8.3, 3)),
     type = "l", xlab = "Temperature", ylab = "Tolerance", lwd = 2)
lines(norm_fun(x = 0:25, mu = 15, va = 9)/ max(norm_fun(0:25, 15, 9)),
      type = "l", col = "blue", lwd = 2)
lines(norm_fun(x = 0:25, mu = 17, va = 7)/ max(norm_fun(0:25, 17, 7)),
      type = "l", col = "green", lwd = 2)

legend(x = 2, y = 0.9, legend = c("spp1", "spp2"), lwd = 2, col = c("black", "blue"))

# plot_spatiotemp_hab(hab = hab, moveCov = moveCov, spwn_wk = list("spp1" = 15:18, "spp2" = 15:18), plot.file =  "testfolder")
# 
# 
# 
# #to plot just the temp preferences over time
# source("R/BENS_plot_spatiotemp_hab_justtemp.R")
# 
# BENS_plot_spatiotemp_hab_justtemp(hab = hab, moveCov = moveCov, spwn_wk = list("spp1" = 16:18, "spp2" = 16:19,"spp3" = 16:18, "spp4" = 18:20), plot.file =  "testfolder")
#

@@

```{r fleets}
#initial settings
# fleets <- init_fleet(sim_init = sim, VPT = list("spp1" = 4, "spp2" = 3),
# 		     Qs = list("fleet 1" = c("spp1" = 1e-5, "spp2" = 3e-5),
# 			       "fleet 2" = c("spp1" = 5e-5, "spp2" = 1e-5)
# 			       ),
# 		     fuelC = list("fleet1" = 3, "fleet 2" = 8),
# 		     step_params = list("fleet 1" = c("rate" = 3, "B1" = 1, "B2" = 2, "B3" = 3),
# 					"fleet 2" = c("rate" = 3, "B1" = 2, "B2" = 4, "B3" = 4)
# 					),				
# 		     past_knowledge = TRUE,
# 		     past_year_month = TRUE,
# 		     past_trip = TRUE,
# 		     threshold = 0.7
# 		     )

@@

test_step(step_params = fleets$fleet_params[[1]]$step_params, rev.max = 1e2)
test_step(step_params = fleets$fleet_params[[2]]$step_params, rev.max = 1e2)

## Spatial closure

We set up a spatial closure. There are multiple options in defining
this, but we simply define a static fixed site closure for demonstration
purposes.

```{r close}
# #practice creating a larger data.frame than just single points
# library(tidyr)
# 
# #set x and y min/max which are coordinates on the grid
# xmin <- 6
# xmax <- 10
# ymin <- 26
# ymax <- 40
# 
# 
# x <- xmin:xmax
# y<- ymin:ymax
# 
# #View(crossing(x,y))
# 
# closure <- init_closure(input_coords = data.frame(x = x, y = y),
# 			spp1 = "spp1", year_start = 2)

## Survey

Its also possible to define a survey design using the init_survey function, but
we do not do so for this demonstration. Please refer to the function help file
if this is required.

```{r survey}

source("R/BENS_init_survey.R")

#CURRENTLY NEED TO MAKE SURE THAT N_STATIONS*#YEARS / #STRATA IS A WHOLE NUMBER OTHERWISE DAY, TOW, YEAR WONT LINEUP WITH NUMBER OF STATIONS
#ALSO NEED N_STATION TO BE DIVISIBLE BY STATIONS_PER_DAY
#ALSO NEED N_STATIONS / STATIONS_PER_DAY <= 52 otherwise wont get to all of them in a year results in NA in the matrix

#setup catch log
surv_random <- BENS_init_survey(sim_init = sim,design = 'random_station', n_stations = 80, #this is total per year (20 in each of 4 strata)
                                start_day = 1, stations_per_day = 1, Qs = c("spp1" = 0.1, "spp2"= 0.2),
                                strata_coords = hab$strata, strata_num = hab$stratas )

@@

    source("R/RcppExports_Bens.R")

    #to source a new move_population where I edited to skips most things:
  #1: load file
  Rcpp::sourceCpp("src/Movement_Bens.cpp") #my edited version that skips most things
  #2: allow the function to call other hidden functions from mixfishsim 
  environment(move_population_Bens) <- asNamespace('MixFishSim')
  #3: replace move_population with go_fish_Bens in the MixFishSim package
  assignInNamespace("move_population", move_population_Bens, ns = "MixFishSim")

@@

@@
